### 2019-03-12 manuscript/discussion.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

% outline
% - disentangling effects
% - general value of approach
%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
%
% - our study key findings and details
% - effects of Ploidy (as %R)
% - trumped by BS
% - no difference on background of SC 
% - discuss data/models and their long/shortcomings: 
%	ClaSSE, will's idea for stochastic mapping
%	irreversibility SI -> SC assumption
% - how to improve / what next
%  (may be useful to study in SC families)
%
% - pathways
% 
%- diploidization
% - rate, effect
% - history in solanaceae
%
% - summary

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by hidden states and or/ breeding system when added to models.  % E: is "trumped" too much like gloating? - R-Haha
%R- I added hidden states because that hasn't been discussed ever in polyploidy
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
%B: should we mention here that the hidden states presage these results in each case? Maybe we don't want to make too much of that, but it is fairly interesting. Also possibly hit that in your paragraph(s), below.
In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against diploid genomic background. %B: "in diploids" may work better?
%B: need help with pathway single-sentence summary
Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined support for diploidization in the history of the family, which appears to be weak. %B: "mixed" "not strong" what's the right word here?  E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of complex hypotheses regarding a heterogenous diversification process.
Below, we discuss these key findings and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

@@

>>>>

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.  
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
%Inference of transition rates can appreciably across 
Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family. %B: "mixed" "not strong" what's the right word here?  E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
Below, we discuss these key findings and detail their sometimes confusing and complex implications.
We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

@@

 %%%%%%%%%%%
 %% Notes to Self %%
 %%%%%%%%%%%

% outline
% - disentangling effects
% - general value of approach
%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
%
% - our study key findings and details
% - effects of Ploidy (as %R)
% - trumped by BS
% - no difference on background of SC 
% - discuss data/models and their long/shortcomings: 
%	ClaSSE, will's idea for stochastic mapping
%	irreversibility SI -> SC assumption
% - how to improve / what next
%  (may be useful to study in SC families)
%
% - pathways
% 
%- diploidization
% - rate, effect
% - history in solanaceae
%
% - summary
%
% B: NEW REORG COMMENTS:

% First, include all straightforward Results and related observations
%
% Then try to square the oval of history of DP and IC. Specifically, it is difficult to know what to make of the delta models. Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data. This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head. 
%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong. Why? It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots! Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences. Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha! Otherwise, something is off with the inference delta >> rho in some cases. If so, we'll have to explain paleo-WGD inference, etc. Either way, we fundamentally misunderstand the evolution of one of these two traits.

% including delta makes for some nonsense: ancestral states are all P with rampant recent transitions to D (e.g. Fig. S4/S8. Not backed by any data, including synteny, S-allele trans-polymorphism, or anecdotally bearing experimental data about rates of spontaneous polyploidization and diploidization \citep{husband_2012}.

% PATHWAYS -- mutational opportunity vs lineage transition rate reveals something about selection 
%
% Starting with SI-D, selection favoring establishment of SC-D seems more common than that of SC-P.
% This is because the mutation rates are likely higher to SC-P, but rate 3x lower. "
%Husband2012: The prevalence of polyploidy within plants has led to debate about the influence of genome duplication on net species diversification (Otto and Whitton 2000). Researchers have argued, on one hand, that polyploidy may simply accumulate as a unidirectional mutation with little influence on the rate or trajectory of evolutionary divergence (Stebbins 1971; Meyers and Levin 2006). This idea is based on the premise that ploidy is more likely to increase than decrease in time. Although reversals do occur, the resulting haploids are rela- tively rare and often of reduced fitness and high sterility (Stebbins 1980; Ramsey and Schemske 2002). Alternatively, polyploidy itself may enhance evolutionary diversification leading to increased divergence and species richness (Otto and Whitton 2000). Elevated diversification may occur if polyploidy accelerates adaptive divergence and evolution of reproductive isolation (Otto and Whitton 2000) or reduces species extinction. 
%The presence of intraspecific ploidy diversity has three main causes. First, genome duplication is an ongoing genomic mutation within many natural populations and ploidy polymorphism is therefore expected when poly- ploid formation is common. In fact, Ramsey and Schemske (1998) estimated that polyploids form at the rate of 10 5 per individual per generation, well above the genic mutation rate, and in some cases it may be even higher than that (Husband 2004). Second, depending on the frequency and pathways of polyploid formation, diploids and polyploids may be maintained as genetically homogeneous populations, which are justifiably treated as single taxonomic species. Finally, ploidy polymorphism may arise as an artefact of our imperfect taxonomic approach. Since morphological divergence is often used as a criterion for species delinea- tion, polyploid lineages may not be recognized at the species level, especially in autopolyploids, which often have a close resemblance to their diploid progenitors (Soltis et al. 2007).

### 2019-03-12 manuscript/introduction.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

%NOTE: please separate each line with a line break!

%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.  E: nice!

Species accumulate across the tree of life at different rates. 
A possible explanation for this phenomenon is that species possess various traits or character states that differentially affect rates of diversification. 
Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}. %is it only a "statistical" problem?
In part, the difficulties arise because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one trait with diversification patterns can be misleading.

Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
WGDs have the potential to affect many phentoypes and a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
Polyploids have been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDs have played an important macroevolutionary role driving innovation and diversification in plants.
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
% Testing for the presence of diploidization in the context of diversification is important because the presence of a positive net diversification rate for diploids can be the result of an ancestrally polyploid lineage.
% Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process. %B: I think this should read "...polyploidy is no longer be considered an evolutionary dead end."  %R- This last phrase is from Itay and I rather like it. <-- %B: Hmm. But how would it be considered an "initiator of the diversification process?" It seems to be associated with lower diversification. In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
% E: I wrote the two sentences below to try to reconcile the text and comments below.  But then I realized we don't actually test this hypothesis.
% One hypothesis is that diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
% Testing this hypothesis requires incorporating diploidization in analyses of diversification.
% E: so try again -- okay?
This could affect interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}. % Gordon: Really? I find this hard to believe... %B:  sigh.
Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates.

Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
Polyploidy may facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
% Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.  E: reworded because no one would think the only thing P does is follow BS?
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date.
There is thus a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. 
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI. That's not a concern, I think, they are all "SI systems." And Dioecy etc. is not classically considered a "breeding system", but a "sexual system". That may be a part of the confusion. It's really not Gordon's problem, and we could do something to yet again re-define these terms? Emma?
% E: I vote for using terms correctly without providing a glossary---lead by example?

Given that changes in ploidy and breeding systems are causally related, and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
First, it may be of interest to establish their joint influence on speciation and extinction rates.
If the effects of ploidy and breeding system on lineage diversification are not additive, studying each trait separately may not reveal the effects of their combinations.
Second, the two traits may affect each other's state transitions.
For example, losses of SI in diploids could be caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without an increase in ploidy, and then a subsequent polyploidization. 
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
It remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.

Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction. 
Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates \citep{rabosky_2017}.
%B: This may be a good portion of Discussion--faults and qualities of methods?
%HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.  
%Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately. 
%Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
%However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
%Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates. 
A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates. 
%These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.

Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
First, we present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
Second, we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
Third, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
For all of the ploidy models, we investigate the potential effects of including diploidization.
Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.

>>>>

Species accumulate across the tree of life at different rates. 
A possible explanation for this phenomenon is that species possess various traits or character states that differentially affect rates of diversification. 
Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}. 
Problems arise at least in part because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one trait with diversification patterns can be misleading.

Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
Increase in ploidy can alter many traits and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
%
Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants. 
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
This could affect the interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates, and may exhibit complex interactions with other traits.

Ploidy, for instance, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
Whole genome duplication may help facilitate the loss of SI and the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date.
There is consequently a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.

%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
Given that changes in ploidy and breeding systems are causally related, and may have profound effects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
If the effects of ploidy and breeding system on lineage diversification are not additive, studying each trait separately may not reveal the effects of their combinations.
As a result, it may be of interest to establish their joint influence on speciation and extinction rates.
The two traits may also affect each other's state transitions.
For example, losses of SI in diploids could be caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without an increase in ploidy, and then a subsequent polyploidization. 
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
It remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.

%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
\citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction. 
Soon thereafter, \citet{beaulieu_2016} proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
Because variation in speciation and extinction rates may not be solely due to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates.
A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates.

Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
First, we present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
Second, we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
Third, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
For all of the models involving ploidy, we investigate the potential effects of including diploidization.
Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.

### 2019-03-12 manuscript/main.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
%
% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
\begin{center}
    \textsc{Interaction between ploidy, breeding system, and lineage diversification}

@@

%--------------------------------------------------
% Main text
%--------------------------------------------------

@@

This work was supported by the National Science Foundation (DEB-1655692 to BI, and DEB-1655478 to EEG) and
the Minnesota Supercomputing Institute (MSI) at the University of Minnesota, which provided computing resources for this project.
%BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?

>>>>

% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phytologist
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
\begin{center}
    \textsc{Interaction between ploidy, breeding system, and lineage diversification}

@@

%--------------------------------------------------
% Main Text
%--------------------------------------------------

@@

This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI).  The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
%BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?

### 2019-03-12 manuscript/methods.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
The number of species in each is shown under the corresponding circles in the top row. % FIXME: Rosana, what are the numbers for CP and ?P ?
These categories were then grouped in a manner appropriate to the states of each model.
For example, there are 34 species that are self-compatible and of unknown ploidy; these are coded as either $D$ or $P$ in the D/P models (uncertain, or consistent with either state), as $C$ in the I/C models, and as either $CD$ or $CP$ in the ID/CD/CP models.
In all cases, species were coded as either $A$ or $B$ in the hidden state models.
The models depicted in the different rows use state spaces that are not comparable with one another.
For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same and are not `lumpable' \citep{tarasov_2019}.
%B: Hmmm. The presentation is economical, true, and I think it is retrospectively clear, and valuable, now that I've been thinking about the paper for a while. But it will be very difficult as a source of clarity for someone reading the paper the first time. The top graphical layout also implicitly communicates the idea that all "I" species must be "D", and that all "P" species must be also "C", which is not intuited easily. (Are they encoded like that?) Moreover, it's not made clear that the 'species' in question are, in fact, often polymorphic, which makes DPIC possible. The subsequent models require a bit of thought, which is not too bad, but the treatment of the state coding and specifically the sentence "In all cases, species were coded as either A or B in the hidden state models." is a bit terse for someone not familiar with HiSSE. 
% E: Yes, all I species are D and all P species are C (no SI polyploids).  I don't think there are polymorphic species because species with some SI and some SC pops are called I, and the most common ploidy level was used.
}
\label{figure:stateclassifications}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used for each of the models.
Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
The number of species in each is shown under the corresponding circles in the top row.
These categories were then grouped in a manner appropriate to the states of each model.
For example, there are 34 species that are self-compatible ($C$) with unknown ploidy, and 40 species known to be polyploid($P$), but without breeding system data; the latter can be inferred, because all polyploid species in the dataset are self-compatible (see \citealt{robertson_2011}, and text). %B: I added stuff here
In the D/P models, these are coded as either $D$ or $P$ (uncertain/consistent with either state); in the I/C  models, they are coded as $C$; and in the ID/CD/CP models, they are coded as either $CD$ or $CP$.
In the hidden state models, species were coded as either $A$ or $B$.
The model-specific coding is shown in different rows. 
%B: I think that these two sentences are more appropriate for the model comparison tables and Methods text. If we are admonished to cut text, they could go.
They each use state spaces that are not directly comparable with one another.
For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same, and are therefore not `lumpable' \citep{tarasov_2019}.
%B: [This] will be very difficult as a source of clarity for someone reading the paper the first time.
%E: Yes, all I species are D and all P species are C (no SI polyploids).  I don't think there are polymorphic species because species with some SI and some SC pops are called I, and the most common ploidy level was used.
}
\label{figure:stateclassifications}
\end{figure}

### 2019-03-12 manuscript/results.tex
removed some tex warnings
- backup of work on manuscript--all sections; extensive comment deletions and edits.

We considered models both with or without diploidization in order to explore its effects on the estimates of state-dependent diversification.
For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).

% R will add a bit about param confidence

@@

\begin{figure}
\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
\caption{Net diversification rates for all models that exclude diploidization.}
\label{figure:netdivall}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
\caption{Net diversification rates for all models that include diploidization.}  
\label{figure:netdivnodip}
\end{figure}

@@

\begin{suppfigure}
\includegraphics[width=\textwidth]{asrDP.pdf}
\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
\label{suppfigure:DPnodip}
\end{suppfigure}

>>>>

We considered models both with or without diploidization in order to explore its effects on the estimates of state-dependent diversification.
For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
% R will add a bit about param confidence

@@

%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
\begin{figure}
\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
\caption{Net diversification rates for all models without diploidization rate estimate. 
Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
\label{figure:netdivall}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
\caption{Net diversification rates for all models that include diploidization ($\delta$). 
Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}  
\label{figure:netdivnodip}
\end{figure}

@@

\begin{suppfigure}
\includegraphics[width=\textwidth]{asrDP.pdf}
\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
\label{suppfigure:DPnodipasr}
\end{suppfigure}
